---
title: "Dataset on Suicide Attacks"
author: "Leah Shiferaw"
date: "2024-03-24"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(readxl)
library(patchwork)


dsat = "dsat_dist_2020_10.xlsx"
df = excel_sheets(dsat) |>
  set_names() |>
  map_dfr(~ read_excel(dsat, sheet = .x), .id = "dsat")

df2 = df |>
  select(date_year, weapon_txt, admin0_txt, group_name)

```

## Overview of the data

The Database on Suicide Attacks (DSAT) was created by researchers at the 
University of Chicago's Project on Security and Threats (CPOST). The data set 
examines worldwide suicide attacks from 1982 to 2019. 

In this project, a suicide attack is defined as "an attack in which an attacker kills himself in a deliberate attack to kill others." Only suicide attacks committed by non-state actors are included in the data set.


This presentation will focus on the following variables:

- Group
- Country
- Year
- Weapon


## Overview of the data (cont.)

```{r top groups}

top_groups = df2 |>
  filter(!is.na(group_name)) |>
  count(group_name, name = "attack_count") |>
  arrange(desc(attack_count)) |>
  head(5)

kable(top_groups, caption = "Top Five Groups")
knitr::kable(top_groups, caption = "Top Five Groups")

```
```{r top countries}

top_countries = df2 |>
  filter(!is.na(admin0_txt)) |>
  count(admin0_txt, name = "attack_location") |>
  arrange(desc(attack_location)) |>
  head(5)

kable(top_countries, caption = "Top Five Countries")
knitr::kable(top_countries, caption = "Top Five Countries")

```

## Analysis 1: Suicide attacks by year


```{r}

summary(df2$date_year)

df3 = df2 |>
  count(date_year) |>
  rename(n = n)


ggplot(df3, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 2) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks by Year") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

date1 = subset(df3, date_year >= 1982 & date_year <= 2000)

plot1 = ggplot(date1, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 2) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks: 1982 to 2000") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

date2 = subset(df3, date_year >= 2001 & date_year <= 2019)

plot2 = ggplot(date2, aes(x = date_year, y = n)) +
  geom_line(color = "#0069c0") +
  geom_point(color = "black", size = 2) +
  labs(x = "Year", y = "Count", title = "Suicide Attacks: 2001 to 2019") +
  scale_y_continuous(limits = c(0, 1000)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
print(plot1)
print(plot2)

combined_plots <- plot1 + plot2
combined_plots
print(combined_plots)

```

## Analysis 2: Most common types of weapons

The researchers have identified 17 types of weapons used to carry out suicide attacks. The top five most frequently used weapons were:

- Car bombs
- Belt bombs
- Truck bombs
- Unspecified
- Motorcycle bombs


```{r weapon types}

weapons = df2 |>
  count(weapon_txt) |>
  arrange(desc(n)) |> 
  filter(!is.na(weapon_txt))
  
top_weapons = head(weapons, 5)  

ggplot(top_weapons, aes(x = reorder(weapon_txt, n), y = n, fill = weapon_txt)) +
  geom_bar(stat = "identity") +
  labs(x = "Weapon Type", y = "Count", title = "Top Five Most Frequent Weapon Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Reds") 
  

```

## Weapon Types by Year


```{r top 5 weapons}

weapon_counts = df2 |>
  count(date_year, weapon_txt) |>
  arrange(date_year, desc(n)) |>
  group_by(date_year) |>
  slice_head(n = 5) |>
  ungroup()

ggplot(weapon_counts, aes(x = date_year, y = n, color = weapon_txt)) +
  geom_line() +
  labs(x = "Year", y = "Count", title = "Top Five Weapons by Year") +
  theme_minimal() +
  theme(legend.position = "bottom")

selected_weapons = c("Car bomb", "Belt bomb", "Truck bomb", "Unspecified", "Motorcycle bomb")


filtered_df = df2 |>
  filter(weapon_txt %in% selected_weapons)

weapon_counts = filtered_df |>
  count(date_year, weapon_txt) |>
  arrange(date_year, desc(n)) |>
  group_by(date_year) |>
  slice_head(n = 5) |>
  ungroup()

ggplot(weapon_counts, aes(x = date_year, y = n, color = weapon_txt)) +
  geom_line() +
  labs(x = "Year", y = "Count", title = "Top Five Weapons by Year") +
  theme_minimal() +
  theme(legend.position = "bottom")

print(ggplot)

```



